---
title: "R Notebook"
output: html_notebook
---

Example from [Ron Yurko's blog](http://www.stat.cmu.edu/~ryurko/post/bayesian-baby-steps-intro/)

Notes / Comments:

- uses some nfl stat scraping tool that he built. Provides a nice data set to play with for the NFL
- Not actually doing the Bayesian update here in the sense that I was hoping for; he's just using the cumulative stats with the analytic result. He says something about this in the blog post but it's not what I was hoping to see, which is whether he's doign something clever to use the posterior for one game as the prior for the next.

```{r setup}
library(tidyverse)
library(nflscrapR)

juju_games_raw <- read_csv("https://raw.githubusercontent.com/ryurko/nflscrapR-data/master/legacy_data/season_play_by_play/pbp_2017.csv")
```

```{r}
juju_games <- juju_games_raw %>% 
  # Filter down only to the pass attempts to JuJu based on his GSIS ID 00-0033857:
  filter(Receiver_ID == "00-0033857",
         PassAttempt == 1) %>%
  # Only select the GameID and PassOutcome columns:
  select(GameID, PassOutcome) %>%
  # Calculate the number of receptions, targets, and catch rate in each game:
  group_by(GameID) %>%
  summarise(receptions = length(which(PassOutcome == "Complete")),
            targets = n(),
            catch_rate = receptions / targets) %>%
  # Calculate cumulative stats:
  mutate(total_receptions = cumsum(receptions),
         total_targets = cumsum(targets),
         total_catch_rate = total_receptions / total_targets,
         # Columns to be used later:
         index = 1:n(),
         game_index = paste("game_", index, sep = ""),
         game_index = fct_relevel(factor(game_index),
                                  "game_1", "game_2", "game_3",
                                  "game_4", "game_5", "game_6",
                                  "game_7", "game_8", "game_9",
                                  "game_10", "game_11", "game_12",
                                  "game_13"))
```

```{r}
p_grid <- seq(from = 0, to = 1, by = .01)
prior <- rep(1, 101)
likelihood <- dbinom(x = juju_games$receptions[1],
                     size = juju_games$targets[1],
                     prob = p_grid)
bayes_numerator <- likelihood * prior
posterior <- bayes_numerator / sum(bayes_numerator)
```


```{r fig.height=6, fig.width=12}
# Create a data frame by applying the grid approximation steps to each row
# of juju_games:
game_posteriors <- map_dfc(c(1:nrow(juju_games)),
                           function(x) {
                             p_grid <- seq(from = 0, to = 1, by = .01)
                             prior <- rep(1, 101)
                             likelihood <- dbinom(x = juju_games$total_receptions[x],
                                                  size = juju_games$total_targets[x],
                                                  prob = p_grid)
                             bayes_numerator <- likelihood * prior
                             posterior <- bayes_numerator / sum(bayes_numerator)
                             # Return this as a data frame:
                             result <- data.frame(posterior)
                             colnames(result) <- paste("game_", x, sep = "")
                             return(result)
                           })

# Join these columns with p_grid and column for the prior probability:
data.frame(p_grid = p_grid, prior = rep(1 / 101, 101)) %>%
  bind_cols(game_posteriors) %>% 
  # Gather the columns so the data is long, one row for each week and grid value
  gather(key = "game_index", value = "posterior_prob", -p_grid) %>%
  # Relevel the game_index variable:
  mutate(game_index = fct_relevel(factor(game_index),
                                  "prior", "game_1", "game_2", "game_3",
                                  "game_4", "game_5", "game_6",
                                  "game_7", "game_8", "game_9",
                                  "game_10", "game_11", "game_12",
                                  "game_13")) %>%
  # Visualize the posteriors for each game:
  ggplot(aes(x = p_grid, y = posterior_prob)) + 
  geom_line(color = "darkblue") +
  facet_wrap(~ game_index) +
  # Add vertical lines for each cumulative observed rate
  geom_vline(data = juju_games, 
             aes(xintercept = total_catch_rate), color = "darkorange",
             linetype = "dashed", size = 1, alpha = .5) +
  geom_text(data = juju_games, size = 3,
            x = .25, y = .025, aes(label = paste("Caught", 
                                               receptions, "of",
                                               targets, sep = " "))) +
  # Label!
  labs(x = "Catch rate", y = "Posterior probability",
       title = "Posterior approximation for JuJu's catch rate after each game") +
  # Clean it up:
  theme_bw() + theme(axis.text.y = element_text(size = 10), 
                     axis.text.x = element_text(size = 6),
                     title = element_text(size = 10)) +
  ylim(0,0.1)
```

